Self-assembly of colloid-cholesteric composites: A route to switchable optical materials

### 2013-11-21 papers/2012-sandwich/CholColloids.tex
updated ms (21 Nov, davide)
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@2258 40876441-3496-0410-a613-a2fca35b2ae2

An interesting question, relevant for applications to devices, is whether 
these structures can be modified, or switched, with electric fields.
Fig.~3 addresses this question, starting from amorphous Blue Phase III,
this time at a higher chirality. Dispersing colloids in this networks
results in the particle occupying the available nodes, where the order
parameter is low (Fig.~3A; Suppl. Movie 5). Applying a high enough field along the $z$ 
direction leads to a switching to a new phase, where colloidal arrays fit 
within a network with distorted hexagonal cells (Fig.~3B; Suppl. Movie 6). Remarkably, upon 
field removal this honeycomb structure does not find its way back to the 
configuration before switching on the field, but gets stuck into a metastable 
structure (Fig. S7), with residual anisotropy along the field switching 
direction.
The resulting multistability is a prerequisite for new generation 
switchable electrooptic devices for use in energy-saving electronics-free
appliances, smart glass, and e-paper to cite just a few examples.

@@

The surprising wealth of colloid-BP composites obtained in Figs.~1-3 is the complex consequence of a simple competition between two self-assembly principles. 
First, blue phases provide a three-dimensional template onto which colloidal particles are attracted so as to reduce the elastic stresses arising at the disclination cores~\cite{miha}. Second, colloidal intrusions in a liquid crystal raise the free energy locally; sharing this cost creates a generic tendency to aggregate into clusters~\cite{tiffany}. Which of these factors dominates depends on $w$, a dimensionless measure of the anchoring strength of the director field at the colloidal surface. For $w\ll 1$, the templating principle dominates. Colloidal crystals, gels or twisted rings can then arise. For $w \gg 1$, interparticle attractions defeat the templating tendency, and we predict disconnected aggregates, percolating gels, or helical colloidal ropes. In both cases, control is offered by the process route, colloid concentration, and the anchoring conditions at sample walls. We have also shown that electric fields can be used to switch between metastable states, providing a route to future multistable energy-saving devices. 
%This variety of metastable structures, with tunable properties that can combine the mechanical and optical features of colloids and liquid crystals~\cite{lavrentovich}, offers promise for the design of multistable liquid crystal devices, which are at the forefront of current technological developments in this field. 
We have not explored in this work their switchability by flow \cite{flowswitch}, but this could again be inherited from the underlying BP dynamics \cite{adriano,permeation2}, in which case they may also be relevant to the emerging technology of optofluidics~\cite{optofluidics}.

\bigskip
\textbf{Acknowledgements}
KS, OH, and JSL are supported by UK EPSRC grants EP/J007404/1, EP/G036136/1,
and EP/F054750/1, respectively. MEC is supported by the Royal Society.
We thank both Hector and PRACE for computational resources.

>>>>

An interesting question, relevant for applications to devices, is whether 
these structures can be modified, or switched, with electric fields.
Fig.~3 addresses this question, starting from amorphous Blue Phase III,
this time at a higher chirality. Dispersing colloids in this networks
results in the particle occupying the available nodes, where the order
parameter is low (Fig.~3A; Suppl. Movie 5). Applying a high enough field 
along the $x$ direction leads to a switching to a new phase, where colloidal 
arrays fit within a network with distorted hexagonal cells (Fig.~3B 
and Fig. S7; Suppl. Movie 6). 
Remarkably, upon field removal this honeycomb structure does not find its 
way back to the 
configuration before switching on the field, but gets stuck into a metastable 
structure (Fig. S7), with residual anisotropy along the field switching 
direction. Cycling the field along the $x$ and $y$ directions leads to
distinct metastable structures, which are retained after field removal
(Fig. S8 and S9).
%The resulting multistability is a prerequisite for new generation 
%switchable electrooptic devices for use in energy-saving electronics-free
%appliances, smart glass, and e-paper to cite just a few examples.

@@

The surprising wealth of colloid-BP composites obtained in Figs.~1-3 is the complex consequence of a simple competition between two self-assembly principles. 
First, blue phases provide a three-dimensional template onto which colloidal particles are attracted so as to reduce the elastic stresses arising at the disclination cores~\cite{miha}. Second, colloidal intrusions in a liquid crystal raise the free energy locally; sharing this cost creates a generic tendency to aggregate into clusters~\cite{tiffany}. Which of these factors dominates depends on $w$, a dimensionless measure of the anchoring strength of the director field at the colloidal surface. For $w\ll 1$, the templating principle dominates. Colloidal crystals, gels or twisted rings can then arise. For $w \gg 1$, interparticle attractions defeat the templating tendency, and we predict disconnected aggregates, percolating gels, or helical colloidal ropes. In both cases, control is offered by the process route, colloid concentration, and the anchoring conditions at sample walls. We have also shown that electric fields can be used to switch between metastable states, providing a possible route to future multistable device applications.
%This variety of metastable structures, with tunable properties that can combine the mechanical and optical features of colloids and liquid crystals~\cite{lavrentovich}, offers promise for the design of multistable liquid crystal devices, which are at the forefront of current technological developments in this field. 
We have not explored in this work switchability by flow \cite{flowswitch}, but this could again be inherited from the underlying BP dynamics \cite{adriano,permeation2}, in which case the same materials may also be relevant to the emerging technology of optofluidics~\cite{optofluidics}.

\bigskip
\textbf{Acknowledgements}
We acknowledge support by UK EPSRC grants EP/J007404/1 (KS and JSL), 
EP/G036136/1 (OH), and EP/I030298/1 (KS and JSL).
MEC is supported by the Royal Society.
We thank both Hector and PRACE for computational resources.
